Blame and Coercion: Together Again for the Second Time

### 2019-10-04 paper.tex
update ;

Implementing gradually typed languages naively on top of blame calculus 
suffices sever space leak.
Translation to Blame Calculus can wrap some expressions with a cast, including 
those at tail position. Thus, at runtime, mutually tail-recursive functions in 
source language can accumulate casts at tail position.

@@

So far, it seems that the coercion-based approach is the best way to combine 
space-efficiency and blame-tracking: coercion is claimed easy to 
understand \cite{garcia2013calculating}\cite{siek2015blame} and is shown 
easy enough to implement in a compiler \cite{kuhlenschmidt2018efficient}.
Coercion is not satisfactory, however, in at least two aspects.
Firstly, its compose function is not structurally recursive, which makes 
it difficult to study coercion formally.
Many developers of Grift report that it is tricky to convince their 
proof assistants that composition terminates. 
Secondly, it is desirable to understand the nature of ``normalized cast'', but 
the coercion-based approach capture this concept indirectly because normalized 
coercion is defined taking a subset of coercion.

Perhaps unsurprisingly, threesome has a straightforward recursive 
implementation. And its definition is self-standing. Together with 
\citet{garcia2013calculating}'s work, they suggest that there can be a cast 
representation whose definition is self-standing, and whose composition is 
structurally recursive. Super-coercion introduced in 
\citet{garcia2013calculating} could 
be a promising candidate. However, its definition is a bit complicated. It uses 
10 constructors to deal with an elementary type system with only base types and 
function types. And four constructors are directly related to 
function types. Thus, super-coercion might not scale very well to more 
sophisticated type systems.

@@

	\fbox{$ \Gamma \vdash e \vdash T $}
	\begin{gather*}
		\inference{
			S \sim T & \Gamma \vdash e : S 
		}{
			\judgetype{\Gamma}{\langle T \Leftarrow^l S \rangle t}{T}
		} \quad
		\inference{
		}{
			\judgetype{\Gamma}{\mathtt{blame} l}{T}
		}
	\end{gather*}

@@

We describe the dynamic semantics of \lazyD{} Blame Calculus with the CEK 
machine (\citet{felleisen1986control}).

@@

\begin{figure}
	\[ 
	\begin{array}{lclr}
	\stxrule{c}{hyper-coercions}{
		id* \mid{}
		( h , ( b , t ) )
	}
	\stxrule{h}{heads}{
		\epsilon \mid{}
		?^l
	}
	\stxrule{b}{bodies}{
		U \mid{}
		c_1 \rightarrow c_2 \mid{}
		c_1 \times c_2 \mid{}
		c_1 \plus c_2
	}
	\stxrule{t}{tails}{
		\epsilon \mid{}
		! \mid{}
		\bot^l
	}
	\end{array}
	\]

@@

	\fbox{$ b \fatsemi^l (b,t) = (b,t) $}
	\[ 
	\begin{array}{rclclr}
	\comprule{U}{(U,t)}{
		(U,t)
	}{}
	\comprule{c_1 \rightarrow c_2}{(c_3 \rightarrow c_4,t)}{
		(c_3 \fatsemi^{l} c_1 \rightarrow c_2 \fatsemi^l c_4, t)
	}{}
	\comprule{c_1 \times c_2}{(c_3 \times c_4,t)}{
		(c_3 \fatsemi^{l} c_1 \times c_2 \fatsemi^l c_4, t)
	}{}
	\comprule{c_1 \plus c_2}{(c_3 \plus c_4,t)}{
		(c_3 \fatsemi^{l} c_1 \plus c_2 \fatsemi^l c_4, t)
	}{}
	\comprule{b_1}{(b_2,t_2)}{
		(b_1,\bot^l)
	}{}
	\end{array}
	\]

>>>>

Implementing gradually typed languages naively on top of blame calculus 
suffices sever space leak.
For example, translation to Blame Calculus can wrap some expressions with a 
cast, including those at tail position. Thus, at runtime, mutually 
tail-recursive functions in source language can accumulate casts at tail 
position. There are more examples in \citet{herman2010space}.

@@

So far, it seems that the coercion-based approach is the best way to combine 
space-efficiency and blame-tracking: coercion is claimed easy to 
understand \cite{garcia2013calculating}\cite{siek2015blame} and is shown 
easy enough to implement in a compiler \cite{kuhlenschmidt2018efficient}.
Coercion is not satisfactory, however, in at least two aspects.
Firstly, its compose function is not structurally recursive, which makes 
it difficult to study coercion formally.
Many developers of Grift report that it is tricky to convince their 
proof assistants that composition terminates. 
Secondly, it is desirable to understand the nature of ``normalized cast'', but 
the coercion-based approach capture this concept indirectly because normalized 
coercion is defined ny taking a subset of coercion.

Perhaps unsurprisingly, threesome has a structurally recursive compose. And its 
definition is self-standing. Together with \citet{garcia2013calculating}'s 
work, they suggest that there can be a cast representation whose definition is 
self-standing, and whose composition is structurally recursive. 
Super-coercion introduced in \citet{garcia2013calculating} could be a promising 
candidate. However, its definition is a bit complicated. It uses 10 
constructors to deal with an elementary type system with only base types and 
function types. And four constructors are directly related to function types. 
Thus, super-coercion might not scale very well to more sophisticated type 
systems.

@@

	\fbox{$ \Gamma \vdash e \vdash T $}
	\begin{gather*}
		\inference{
			S \sim T & \Gamma \vdash e : S 
		}{
			\judgetype{\Gamma}{\langle T \Leftarrow^l S \rangle t}{T}
		} \quad
		\inference{
		}{
			\judgetype{\Gamma}{\mathtt{blame} \; l}{T}
		}
	\end{gather*}

@@

We describe the dynamic semantics of \lazyD{} Blame Calculus with the CEK 
machine (\citet{felleisen1986control}). We choose CEK machine because to 
achieve space efficiency, we need to change continuations. 
Thus, abstract machines that include continuation in states is quite convenient 
for us. And CEK is one of the simplest machine of this kind.

@@

\begin{figure}
	\[ 
	\begin{array}{lclr}
	\stxrule{c}{hyper-coercions}{
		id* \mid{}
		( h , ( b , t ) )
	}
	\stxrule{h}{heads}{
		\epsilon \mid{}
		?^l
	}
	\stxrule{m}{middles}{
		U \mid{}
		c_1 \rightarrow c_2 \mid{}
		c_1 \times c_2 \mid{}
		c_1 \plus c_2
	}
	\stxrule{t}{tails}{
		\epsilon \mid{}
		! \mid{}
		\bot^l
	}
	\end{array}
	\]

@@

	\fbox{$ b \fatsemi^l (b,t) = (b,t) $}
	\[ 
	\begin{array}{rclclr}
	\comprule{U}{(U,t)}{
		(U,t)
	}{}
	\comprule{c_1 \rightarrow c_2}{(c_3 \rightarrow c_4,t)}{
		(c_3 \fatsemi^{l} c_1 \rightarrow c_2 \fatsemi^l c_4, t)
	}{}
	\comprule{c_1 \times c_2}{(c_3 \times c_4,t)}{
		(c_1 \fatsemi^{l} c_3 \times c_2 \fatsemi^l c_4, t)
	}{}
	\comprule{c_1 \plus c_2}{(c_3 \plus c_4,t)}{
		(c_1 \fatsemi^{l} c_3 \plus c_2 \fatsemi^l c_4, t)
	}{}
	\comprule{b_1}{(b_2,t)}{
		(b_1,\bot^l)
	}{b_1 }
	\end{array}
	\]
